Structured Versus Decorated Cospans

### 2021-06-15 cospans_new.tex
worked on material in and around Thms. 2.1-2.2

In this section we build symmetric monoidal double categories of decorated cospans, and then study the functoriality of this construction.   \cref{thm:decorated_cospans_1} explains how to construct a double category of decorated cospans from a lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat, \times)$  whenever $A$ is a cocartesian.  \cref{thm:decorated_cospans_2} gives conditions under which this double category is symmetric monoidal.   The definition of lax monoidal pseudofunctor is recalled in \cref{subsec:bicats}, and the definition of symmetric monoidal double category is recalled in \cref{defn:double_category}.

In all that follows, when we say a category `has finite colimits' we mean it is equipped with a choice of colimit for every finite diagram.   Thus, if $\A$ has coproducts it gives a cocartesian monoidal category $(\A,+)$: that is, a symmetric monoidal category where the monoidal structure is given by the chosen binary coproducts and initial object.   However, when we say a functor `preserves finite colimits', it need only do this up to canonical isomorphism, unless otherwise specified.

@@

We prove both these theorems using the work of Shulman \cite{Shulman2008}, which we recall here.   The concept of `Beck--Chevalley monoidal bifibration' is explained in \cref{subsec:fibrations}.

@@

\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is a pair of objects $a,b \in \A$ together with an
object $c \in C$ with $\Phi(c) = a + b$.
\item a 2-morphism is a pair of morphisms $f \maps a \to a'$, $g \maps b \to b'$ 
in $\A$ together with a morphism $h \maps c \to c'$ with $\Phi(h) = f + g$.
\end{itemize}
If $\Phi$ is a symmetric monoidal bifibration then $\Fr(\Phi)$ becomes a symmetric
monoidal double category.
\end{lem}

@@

We  prove Theorems \ref{thm:decorated_cospans_1} and \ref{thm:decorated_cospans_2} by applying this lemma to a particular monoidal bifibration.  To get our hands on this starting from our lax monoidal pseudofunctor $F \maps (A,+) \to (\Cat,\times)$, we need a monoidal version of the Grothendieck construction \cite{MV}.   We recall the Grothendieck construction in \ref{subsec:fibrations}; briefly,
this gives a bijection between pseudofunctors $F \maps \A \to \Cat$ and opfibrations
$U \maps \inta F \to \A$.

{\chris It feels like there is some repetition in the material below and the sketch of the proof of 2.4. I believe the previous version of the paper had a briefer text here. No problem if you want to keep as is.}

Moeller and the third author \cite{MV} showed that
similarly, given a (symmetric) monoidal category $\A$, there is a bijection between 
\begin{itemize}
\item \define{(symmetric) monoidal opindexed categories}: (symmetric) lax monoidal pseudofunctors \hfill \break $F\maps (\A, \otimes_A) \to (\Cat, \times)$
\end{itemize}
and 
\begin{itemize}
\item \define{(symmetric) monoidal opfibrations}: opfibrations $U \maps \X \to \A$ where $\X$ and $\A$ are (symmetric) monoidal, $U$ is (symmetric) strict monoidal, and $\otimes_\X$ preserves cocartesian liftings. 
\end{itemize}  
If the monoidal structure on $\A$ is in fact cocartesian, as it is in our case, there is a further correspondence between these structures and ordinary pseudofunctors $F\maps \A \to \MonCat$ (or to $\SMC$ in the symmetric case).

In fact, all these correspondences lift to 2-equivalences between appropriate 2-categories:

\begin{lem}
\label{lem:MonGroth}
There is a 2-equivalence between the 2-categories of monoidal opfibrations and monoidal opindexed categories, and if the base is cocartesian monoidal, there is a 2-equivalence between these and pseudofunctors into $\MonCat$.   Similarly there is a 2-equivalence between symmetric monoidal opfibrations and symmetric monoidal opindexed categories, and if the base is cocartesian monoidal, also a 2-equivalence between these and pseudofunctors into $\SMC$.   
\end{lem}

@@

{\chris Some text saying that we have now set up the necessary framework to use Mike's result and obtain our double categories would be nice here}

\begin{proof}[Proof of \cref{thm:decorated_cospans_1}] 
To obtain the desired double category we apply \cref{lem:shulman} to a bifibration $\Phi \maps \A/U \to \A$ which we now define.   Let $U \maps \inta F \to \A$ be the monoidal opfibration associated to the lax monoidal pseudofunctor $F \maps \A \to \Cat$. {\chris Couldn't we just recall the appendix here?}  Thus:
\begin{itemize}
\item an object of $\inta F$ is a pair $(b,p)$ where $b \in \A$ and $p \in F(b)$;
\item a morphism in $\inta F$ from $(b,p)$ to $(b',p')$ is a pair $(g,\zeta)$
where $g \maps b \to b'$ is a morphism in $\A$ and $\zeta \maps F(f)(p) \to p'$ is
a morphism in $F(b')$.
\end{itemize}
%%% Kenny: The f and g in the second item above need to be the same
The functor $U$ maps objects and morphisms to their first component.   We thus obtain an explicit description of the comma category $\A/U$ {\chris well the arrows are described like that in any slice category, there is nothing special for the Grothendieck category? Perhaps the slice description could be omitted or more brief?}:
\begin{itemize}
\item an object of $\A/U$ is a triple $(a,x,a \xrightarrow{i} U(x))$ with $a \in \A$ and $x=(b,p) \in \inta F$;
\item a morphism in $\A/U$ from $(a,x,a \xrightarrow{i} U(x))$ to $(a',x',a' \xrightarrow{i'} U(x'))$ is a pair $(f,\theta)$ where $f \maps a \to a'$ is a morphism in $\A$ and $\theta \maps x \to x'$ is a morphism in $\inta F$ such that the following square commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$U(x)$};
\node (D) at (1.5,-1) {$U(x')$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge node[right]{$i'$} (D)
(C) edge node[below]{$U(\theta)$} (D);
\end{tikzpicture}
\]
\end{itemize}
The functor $\Phi \maps \A/U \to \A$ is defined on objects by $\Phi(a,x,a \xrightarrow{i} U(x))=a$ and on morphisms by $\Phi(f,\theta)=f$. We have a monoidal structure on $\A/U$ where the tensor product of two objects $(a_1,x_1,a_1 \xrightarrow{i_1} U(x_1))$ and $(a_2,x_2,a_2 \xrightarrow{i_2} U(x_2))$ is given by
\[  (a_1,x_1,a_1 \xrightarrow{i_1} U(x_1)) \otimes (a_2,x_2,a_2 \xrightarrow{i_2} U(x_2)) = (a_1+a_2,x_1 \otimes x_2,a_1+a_2 \xrightarrow{i_1+i_2} U(x_1 \otimes x_2)) \]
where we have left the equality $U(x_1)+U(x_2) = U(x_1 \otimes x_2)$ implicit in the above composite, for reasons of space. The monoidal unit is given by $(0_\A,0_\X,0_\A \to U(0_\X))$. It is easy to see that the functor $\Phi$ is a strict monoidal using this tensor product.

@@

One can also check that the functor $\Phi$ is an opfibration, but this takes a bit more work. {\chris ...or ``is less straightforward'' to express more formally perhaps?} Given a morphism $f \maps a \to a'$ in $\A$ and an object $(a,x,a \xrightarrow{i} U(x)) \in \A/U$ over $a$, we shall show that a cocartesian lifting of $f$ is given by 
\[ (f,\theta) \maps (a,x,a \xrightarrow{i} U(x)) \, \to \, (a',x', a' \xrightarrow{i'} U(x')). \]
%%% Kenny: Just an observation: we're using a morphism f:a-->b for the fibration part and a morphism f:a-->a' for the opfibration part.
Here $i'$ arises from the following pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$U(x)$};
\node (D) at (1.5,-1) {$U(x)+_a a'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge [dashed] node[right]{$i'$} (D)
(C) edge [dashed] node[below]{$g$} (D);
\end{tikzpicture}
\]
The object $x' \in \inta F$ is defined as follows.  Recall that $x \in \inta F$ is a
pair $(b,p)$ with $b\in \A$ and $p \in F(b)$, and note that $U(x) = b$.   Define 
$x'$ to be the pair $(b',p')$ where $b' = U(x)+_a a'$ is the pushout in the above
diagram and $p' = F(g)(p)$.  Then $U(x') = b'$ so we have 
$i' \maps a' \to U(x')$ as desired.  Finally, we define $\theta \maps x \to x'$ to be 
the morphism in $\inta F$ given by the pair $(g,\id)$ where $g \maps b \to b'$ is
given in the above diagram and $\id \maps F(g)(p) \to p'$ is the identity.

@@

Lastly, the monoidal bifibration $\Phi$ satisfies the Beck--Chevalley condition. Given a pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$d$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
we need to show that the following square commutes up to a natural isomorphism:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(\A/U)_a$};
\node (B) at (1.5,0) {$\A/U)_b$};
\node (C) at (0,-1) {$(\A/U)_c$};
\node (D) at (1.5,-1) {$(\A/U)_d$};
\node (E) at (0.75,-0.5) {$\Searrow$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^*$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^*$} (C);
\end{tikzpicture}
\]
Let $(b,x,b \xrightarrow{i} U(x))$ be an object of $(\A/U)_b$ where $x=(b',p) \in \A \times F(b')$ so that $U(x)=b'$. Going left, we precompose with $h \maps a \to b$ to obtain the object $(a,x,a \xrightarrow{i \circ h} U(x))$ of $(\A/U)_a$. Then going down, we push forward along $k$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$U(x)$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$c+_{a}U(x)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i \circ h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\psi$} (D);
\end{tikzpicture}
\]
to obtain the object $(c,z,c \xrightarrow{\psi} U(z))$ of $(\A/U)_c$, where $z \in \A \times F(c+_a U(x))$ is given by $z=(c+_a U(x),q)$ where $q$ is given by
\[ 1 \xrightarrow{p} F(U(x)) \xrightarrow{F(\zeta)} F(0+_0 U(x)) \xrightarrow{F(!_c +_{!_a} 1)} F(c+_a U(x)). \]
%and so by construction, we have that 
%\[  U(z)=U(\bot_c+_{\bot_a}x)=c+_a U(x).\]
%{\john Need to refer to a proposition where we show $U$ preserves pushouts strictly.}
 For the other route, first going down, we push the object $(b,x,b \xrightarrow{i} U(x))$ forward along $g \maps b \to d$ by taking the following pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$b$};
\node (B) at (1.5,0) {$U(x)$};
\node (C) at (0,-1) {$d$};
\node (D) at (1.5,-1) {$d+_{b}U(x)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(A) edge node[left]{$g$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\theta$} (D);
\end{tikzpicture}
\]
which yields the object $(d,z',d \xrightarrow{\theta} U(z'))$ of $(\A/U)_d$ where $z' \in \A \times F(d+_b U(x))$ is given by ${z'=(d+_b U(x),q')}$ where $q'$ is given by 
\[ 1 \xrightarrow{p} F(U(x)) \xrightarrow{F(\zeta)} F(0+_0 U(x)) \xrightarrow{F(!_d +_{!_b} 1)} F(d+_b U(x)). \]
We then precompose with $f \maps c \to d$ to obtain the object $(c,z',c \xrightarrow{\theta \circ f} U(z'))$ of $(\A/U)_c$.

@@

\begin{proof}[Proof of \cref{thm:decorated_cospans_2}] We apply \cref{lem:shulman} to $A/U \to A$: why is this a symmetric monoidal bifibration?
\end{proof}

@@

\subsection{Fibrations}\label{subsec:fibrations}

@@

In a bifibration we have both covariant and contravariant reindexing functors, and in 
fact $f_!$ is left adjoint to $f^*$ \cite[Proposition 3.9]{Shulman2008}.   Using this one can easily show that for any commutative square in $\A$
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$d$};
%\node (E) at (0.5,-0.5) {$\Downarrow\theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
the following square commutes up to a specified natural transformation:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\C_a$};
\node (B) at (1.5,0) {$\C_b$};
\node (C) at (0,-1) {$\C_c$};
\node (D) at (1.5,-1) {$\C_d$};
\node (E) at (0.75,-0.5) {$\Searrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^\ast$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^\ast$} (C);
\end{tikzpicture}
\]
where $\theta$ is built as a composite involving the unit of the adjunction between
$g_!$ and $g^\ast$ and the counit of the adjunction between $k_!$ and $k^\ast$:
\[    k_! h^\ast \To k_! h^\ast g^\ast g_! = k_! k^\ast f^\ast g^! \To f^\ast g_! .\] 
If $\theta$ is a natural isomorphism whenever the original square in $\A$ is
a pushout, we shall say $\Phi$ is \define{Beck--Chevalley}.  (Shulman  uses the term `strongly co-BC' \cite[Definition 13.21] {Shulman2008}.)

>>>>

In this section we build symmetric monoidal double categories of decorated cospans, and then study the functoriality of this construction.   \cref{thm:decorated_cospans_1} explains how to construct a double category of decorated cospans from a lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat, \times)$  whenever $A$ is a cocartesian.  \cref{thm:decorated_cospans_2} gives conditions under which this double category is symmetric monoidal.   (We recall the definition of lax monoidal pseudofunctor in \cref{subsec:bicats}, and the definition of symmetric monoidal double category in \cref{defn:double_category}.)

In all that follows, when we say a category `has finite colimits' we mean it is equipped with a choice of colimit for every finite diagram.   Thus, if $\A$ has finite colimits it gives a cocartesian monoidal category $(\A,+)$: that is, a symmetric monoidal category where the monoidal structure is given by the chosen binary coproducts and initial object.   However, when we say a functor `preserves finite colimits', it need only do this up to canonical isomorphism, unless otherwise specified.

@@

We prove both these theorems using the work of Shulman \cite{Shulman2008}.  Shulman gives a general way to construct double categories from `Beck--Chevalley monoidal bifibrations'---a concept we explain in \cref{subsec:fibrations}.

@@

\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is a pair of objects $a,b \in \A$ together with an
object $c \in \C$ with $\Phi(c) = a + b$.
\item a 2-morphism is a pair of morphisms $f \maps a \to a'$, $g \maps b \to b'$ 
in $\A$ together with a morphism $h \maps c \to c'$ with $\Phi(h) = f + g$.
\end{itemize}
If $\Phi$ is a Beck--Chevalley symmetric monoidal bifibration then $\Fr(\Phi)$ becomes a symmetric
monoidal double category.
\end{lem}

@@

Note that when $\C$ is the arrow category of $\A$ and $\Phi$ maps any arrow in $\A$ to its
domain, an object of $\Fr(\Phi)$ simply amounts to a cospan in $\C$.  Thus, in this case $\Fr(\Phi)$ is the double category of cospans in $\C$.    We shall use variations on this idea to define double categories of decorated and structured cospans, and also prove that under certain conditions these double categories are equivalent.

We start with decorated cospans, proving Theorems \ref{thm:decorated_cospans_1} and \ref{thm:decorated_cospans_2} by applying Shulman's result to a particular monoidal bifibration built
from a lax monoidal pseudofunctor $F \maps (A,+) \to (\Cat,\times)$.    The Grothendieck construction gives a bijection between pseudofunctors $F \maps \A \to \Cat$ and opfibrations $U \maps \inta F \to \A$.  We need two refinements of this construction: one that gives monoidal obfibrations, and one that gives symmetric monoidal opfibrations.  We explain the Grothendieck construction and all these related concepts in \cref{subsec:fibrations}; here we simply recall the result we need.

\begin{lem}
\label{lem:MonGroth}
There is a 2-equivalence between the 2-categories of monoidal opfibrations and monoidal opindexed categories with a given base, and if the base is cocartesian monoidal, there is a 2-equivalence between these and pseudofunctors into $\MonCat$.   Similarly there is a 2-equivalence between symmetric monoidal opfibrations and symmetric monoidal opindexed categories, and if the base is cocartesian monoidal, also a 2-equivalence between these and pseudofunctors into $\SMC$.   
\end{lem}

@@

Now we are ready to use Shulman's result to prove  \cref{thm:decorated_cospans_1} and  \cref{thm:decorated_cospans_2}.

\begin{proof}[Proof of \cref{thm:decorated_cospans_1}] 
To obtain the desired double category we apply \cref{lem:shulman} to a bifibration $\Phi \maps \A/U \to \A$ which we now define.   Let $U \maps \inta F \to \A$ be the monoidal opfibration associated to the lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat, \times)$.  As in \cref{subsec:fibrations},
\begin{itemize}
\item an object of $\inta F$ is a pair $(b,p)$ where $b \in \A$ and $p \in F(b)$;
\item a morphism in $\inta F$ from $(b,p)$ to $(b',p')$ is a pair $(f,\zeta)$
where $g \maps b \to b'$ is a morphism in $\A$ and $\zeta \maps F(g)(p) \to p'$ is
a morphism in $F(b')$.
\end{itemize}
The functor $U$ maps objects and morphisms to their first component, so in the comma
category $\A/U$, 
\begin{itemize}
\item an object is a triple $(a,x,a \xrightarrow{i} U(x))$ with $a \in \A$ and $x=(b,p) \in \inta F$;
\item a morphism from $(a,x,a \xrightarrow{i} U(x))$ to $(a',x',a' \xrightarrow{i'} U(x'))$ is a pair $(f,\theta)$ where $f \maps a \to a'$ is a morphism in $\A$ and $\theta \maps x \to x'$ is a morphism in $\inta F$ such that the following square commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$U(x)$};
\node (D) at (1.5,-1) {$U(x')$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge node[right]{$i'$} (D)
(C) edge node[below]{$U(\theta)$} (D);
\end{tikzpicture}
\]
\end{itemize}
The functor $\Phi \maps \A/U \to \A$ is defined on objects by $\Phi(a,x,a \xrightarrow{i} U(x))=a$ and on morphisms by $\Phi(f,\theta)=f$. We have a monoidal structure on $\A/U$ where the tensor product of two objects $(a_1,x_1,a_1 \xrightarrow{i_1} U(x_1))$ and $(a_2,x_2,a_2 \xrightarrow{i_2} U(x_2))$ is given by
\[  (a_1,x_1,a_1 \xrightarrow{i_1} U(x_1)) \otimes (a_2,x_2,a_2 \xrightarrow{i_2} U(x_2)) = (a_1+a_2,x_1 \otimes x_2,a_1+a_2 \xrightarrow{i_1+i_2} U(x_1 \otimes x_2)) \]
where we have left the equality $U(x_1)+U(x_2) = U(x_1 \otimes x_2)$ implicit in the above composite, for reasons of space. The monoidal unit is given by $(0_\A,0_\X,0_\A \to U(0_\X))$. It is easy to see that the functor $\Phi$ is a strict monoidal using this tensor product.

@@

One can also check that the functor $\Phi$ is an opfibration, but this takes a bit more work. Given a morphism $f \maps a \to a'$ in $\A$ and an object $(a,x,a \xrightarrow{i} U(x)) \in \A/U$ over $a$, we shall show that a cocartesian lifting of $f$ is given by 
\[ (f,\theta) \maps (a,x,a \xrightarrow{i} U(x)) \, \to \, (a',x', a' \xrightarrow{i'} U(x')). \]
%%% Kenny: Just an observation: we're using a morphism f:a-->b for the fibration part and a morphism f:a-->a' for the opfibration part.
Here $i'$ arises from the following pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$U(x)$};
\node (D) at (1.5,-1) {$U(x)+_a a'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge [dashed] node[right]{$i'$} (D)
(C) edge [dashed] node[below]{$g$} (D);
\end{tikzpicture}
\]
The object $x' \in \inta F$ is defined as follows.  Recall that $x \in \inta F$ is a
pair $(b,p)$ with $b\in \A$ and $p \in F(b)$, and note that $U(x) = b$.   Define 
$x'$ to be the pair $(b',p')$ where $b' = U(x)+_a a'$ is the pushout in the above
diagram and $p' = F(g)(p)$.  Then $U(x') = b'$ so we have 
$i' \maps a' \to U(x')$ as desired.  Finally, we define $\theta \maps x \to x'$ to be 
the morphism in $\inta F$ given by the pair $(g,\id)$ where $g \maps b \to b'$ is
given in the above diagram and $\id \maps F(g)(p) \to p'$ is the identity.

@@

Lastly, the monoidal bifibration $\Phi$ satisfies the Beck--Chevalley condition. Given a pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$d$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
we need to show that the following square commutes up to a natural isomorphism:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(\A/U)_a$};
\node (B) at (1.5,0) {$(\A/U)_b$};
\node (C) at (0,-1) {$(\A/U)_c$};
\node (D) at (1.5,-1) {$(\A/U)_d$};
\node (E) at (0.75,-0.5) {$\Searrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^*$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^*$} (C);
\end{tikzpicture}
\]
Let $(b,x,b \xrightarrow{i} U(x))$ be an object of $(\A/U)_b$ where $x=(b',p) \in \A \times F(b')$ so that $U(x)=b'$. Going left, we precompose with $h \maps a \to b$ to obtain the object $(a,x,a \xrightarrow{i \circ h} U(x))$ of $(\A/U)_a$. Then going down, we push forward along $k$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$U(x)$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$c+_{a}U(x)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i \circ h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\psi$} (D);
\end{tikzpicture}
\]
to obtain the object $(c,z,c \xrightarrow{\psi} U(z))$ of $(\A/U)_c$, where $z \in \A \times F(c+_a U(x))$ is given by $z=(c+_a U(x),q)$ where $q$ is given by
\[ 1 \xrightarrow{p} F(U(x)) \xrightarrow{F(\zeta)} F(0+_0 U(x)) \xrightarrow{F(!_c +_{!_a} 1)} F(c+_a U(x)). \]
%and so by construction, we have that 
%\[  U(z)=U(\bot_c+_{\bot_a}x)=c+_a U(x).\]
%{\john Need to refer to a proposition where we show $U$ preserves pushouts strictly.}
 For the other route, first going down, we push the object $(b,x,b \xrightarrow{i} U(x))$ forward along $g \maps b \to d$ by taking the following pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$b$};
\node (B) at (1.5,0) {$U(x)$};
\node (C) at (0,-1) {$d$};
\node (D) at (1.5,-1) {$d+_{b}U(x)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(A) edge node[left]{$g$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\theta$} (D);
\end{tikzpicture}
\]
which yields the object $(d,z',d \xrightarrow{\theta} U(z'))$ of $(\A/U)_d$ where $z' \in \A \times F(d+_b U(x))$ is given by ${z'=(d+_b U(x),q')}$ where $q'$ is given by 
\[ 1 \xrightarrow{p} F(U(x)) \xrightarrow{F(\zeta)} F(0+_0 U(x)) \xrightarrow{F(!_d +_{!_b} 1)} F(d+_b U(x)). \]
We then precompose with $f \maps c \to d$ to obtain the object $(c,z',c \xrightarrow{\theta \circ f} U(z'))$ of $(\A/U)_c$.

@@

\begin{proof}[Proof of \cref{thm:decorated_cospans_2}] 
When the lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat, \times)$ of \cref{thm:decorated_cospans_1} is symmetric, the monoidal bifibration $\Phi \maps \A/U \to A$
is symmetric, so \cref{lem:shulman} implies that the double category $F\Csp$ becomes symmetric
monoidal as described.
\end{proof}

@@

\subsection{Fibrations and opfibrations}\label{subsec:fibrations}

@@

In a bifibration we have both covariant and contravariant reindexing functors, and in 
fact $f_!$ is left adjoint to $f^*$ \cite[Proposition 3.9]{Shulman2008}.   Using this one can easily show that for any commutative square in $\A$
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$d$};
%\node (E) at (0.5,-0.5) {$\Downarrow\theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
the following square commutes up to a specified natural transformation:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\C_a$};
\node (B) at (1.5,0) {$\C_b$};
\node (C) at (0,-1) {$\C_c$};
\node (D) at (1.5,-1) {$\C_d$};
\node (E) at (0.75,-0.5) {$\Searrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^\ast$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^\ast$} (C);
\end{tikzpicture}
\]
where $\theta$ is built as a composite involving the unit of the adjunction between
$g_!$ and $g^\ast$ and the counit of the adjunction between $k_!$ and $k^\ast$:
\[    k_! h^\ast \To k_! h^\ast g^\ast g_! = k_! k^\ast f^\ast g^! \To f^\ast g_! .\] 
If $\theta$ is a natural isomorphism whenever the original square in $\A$ is
a pushout, we say $\Phi$ is \define{Beck--Chevalley}.  (Shulman  uses the term `strongly co-BC' \cite[Definition 13.21] {Shulman2008}.)
